Likelihoods for a general class of ARGs under the SMC

### 2025-04-28 paper.tex

Now consider the ``hazard'' associated with recombination.
Begin with an edge $e = (p,c,x,y)$.
This implies that there was a sequence of lineages
from the time of the child $c$ (call this time $t_c$)
back to the time of the parent $p$ (again, $t_p$),
that contained the segment $[x,y)$,
and the lineage that directly inherited from $p$
had the segment $[x,y)$ (see Fig.~\ref{fig:likelihood}).
We therefore know there was no recombination between $x$ and $y$
on any of those lineages
over this time span,
the probability of which is
$\exp(-r \mathcal{A}_e)$, where $r$ is the recombination rate
and $\mathcal{A}_e$ is the total area (eligible links multiplied by length in time) \revpoint{2}{28}
of the edge.
The length in time of the edge is just $t_p - t_c$,
while the number of ``eligible links'' in the edge is the number of adjacent base pairs in the edge
on which recombination did not occur, which is either $y-x$
(if this edge is the leftmost of all edges above node $c$)
or $y-x-1$ (otherwise).
So, the first contribution to the likelihood from this edge $e$ is
\begin{equation}\label{eq:span}
A_e(r) = \exp(-r (y-x - \mathbbm{1}_{x > x_c})(t_p - t_{c})) ,
\end{equation}
where as for lineages, $x_c$ is the leftmost position of the edges whose children are $c$.
% Note that if a recombination did occur, then the number of links not
% affected by recombination is 1 less (see eq.~\eqref{eq:depth}).

>>>>

Now consider the ``hazard'' associated with recombination.
Begin with an edge $e = (p,c,x,y)$.
This implies that there was a sequence of lineages
from the time of the child $c$ (call this time $t_c$)
back to the time of the parent $p$ (again, $t_p$),
that contained the segment $[x,y)$,
and the lineage that directly inherited from $p$
had the segment $[x,y)$ (see Fig.~\ref{fig:likelihood}).
We therefore know there was no recombination between $x$ and $y$
on any of those lineages
over this time span,
the probability of which is
$\exp(-r \mathcal{A}_e)$, where $r$ is the recombination rate
and $\mathcal{A}_e$ is the total ``area'' \revpoint{2}{28}
of the edge.
The area of an edge is its  length in time, $t_p - t_c$,
multiplied by the number of ``eligible links'' along its
span.
The number of eligible links is the number of adjacent base pairs in the edge
on which recombination did not occur, which is either $y-x$
(if this edge is the leftmost of all edges above node $c$)
or $y-x-1$ (otherwise).
So, the first contribution to the likelihood from this edge $e$ is
\begin{equation}\label{eq:span}
A_e(r) = \exp(-r (y-x - \mathbbm{1}_{x > x_c})(t_p - t_{c})) ,
\end{equation}
where as for lineages, $x_c$ is the leftmost position of the edges whose children are $c$.
% Note that if a recombination did occur, then the number of links not
% affected by recombination is 1 less (see eq.~\eqref{eq:depth}).
